EGGs are adhesive!

### 2025-03-12 calco/main.tex
altro lemma

\begin{restatable}{lemma}{pbm}\label{lem:pbmono}
	The class $\pbc$ is closed under composition, decomposition and it is stable under pullbacks and pushouts.
\end{restatable}
\begin{proof}
	Closure under composition and decomposition follows from \Cref{lem:pb1}, while stability under pushouts is the first point of \Cref{prop:kerset}. We are left with stability under pullbacks. Consider the diagrams below, in which the first square is a pullback with $h$ in $\pbc$. In particular, by \Cref{prop:eqhyp_complete}, the two halves of the diagram on the right are pullbacks and so, by \Cref{lem:pb1} the whole rectangle is a pullback too.

	\[\]

		\[
	\xymatrix@C=10pt@R=10pt{&& & &A_1'\ar[dd]|\hole_(.65){a_E}\ar@{>->}[rr]^{h_E'} \ar[dl]_{k_E'} && A_2' \ar[dd]_{b_E} \ar[dl]_{z_E'}  \\\mathcal{P} & & \mathcal{G}_1 &
	A_3'  \ar[dd]_{c_E}\ar@{>->}[rr]^(.7){p_E'} & & A_4' \ar[dd]_(.3){d_E}
	&& V_{\mathcal{P}} && V_1 && Q_1\\
	 &&&&A_1\ar@{>->}[rr]|\hole^(.65){h_E} \ar[dl]^{k_E} && A_2 \ar[dl]^{z_E}  \\
		\mathcal{G}_2 && \mathcal{G}_3 & V_3 \ar@{>->}[rr]_{p_E} & & A_4 && V_2 && V_3 && Q_3 }
	\]

	Thus, by \Cref{lem:pb1} also the right rectangle is a pullback.
\end{proof}

In particular, we are now going to show that $\EqHyp$ is $\reg(\EqHyp)$-adhesive.

@@

	\noindent 
	\parbox{8.5cm}{
	For $Q'_1$, we can make a similar argument. Taking $(S, s_1, s_2)$ as the pullback of $m_2$ along $x_1$, we can use again the composition properties of pullbacks and the second point of \Cref{prop:eqhyp_complete} to guarantee the existence of a monomorphism $m_1\colon Q'_1\to S$ which makes the diagram aside commutative.}\hfill
	\parbox{4cm}{\xymatrix@R=15pt{B'_1\ar@{>>}[r]_{q'_1} \ar@{>->}[d]_{h'_V} \ar@/^.3cm/[rr]^{a_V} & Q'_1 \ar@{>.>}[d]_{m_1} & B_1 \ar@{>>}[dr]^{q_1}\\B'_2 \ar@{>>}[dr]_{q'_2} & S \ar@{>->}[r]^{s_2} \ar[d]_{s_1} & T \ar[r]^{x_2} \ar@{>->}[d]_{x_1}  & Q_1 \ar@{>->}[d]^{h_Q}\\& Q'_2 \ar@{>->}[r]_{m_2}& U \ar[r]_{u_2}& Q_2 }}

@@

We want to extend such a morphism to one of $\EqHyp$ between $\mathcal{G}'_4\to \mathcal{H}$. If we are able to do so we can conclude because uniqueness is guaranteed by \Cref{rem:eqhyp_morphs}.

Now, consider the cube below, $h$ is in $\pbc$ so by \Cref{lem:pb1} we know that its back face is a pullback. We can then apply the second point of \Cref{prop:kerset} to deduce that the square on the right is a pushout.
\[\xymatrix@C=10pt@R=10pt{
	&B_1'\ar@{>>}[dd]|(.53)\hole^(.67){s_2\circ m_1\circ q_1'}\ar@{>->}[rrrr]^{h_V'} \ar[dl]_(.6){k_V'} &&&& B_2' \ar@{>}[dd]^{m_2\circ q_2'} \ar[dl]_(.6){z_V'} & & K_{s_2\circ m_1\circ q'_1}  \ar@{>->}[rr]^-{k_{h'_V}} \ar[dd]_{k_{k'_V}}&& K_{m_2\circ q'_2} \ar[dd]^{k_{z'_V}}\\
	B_3'  \ar[dd]_{m_3\circ q_3'}\ar@{>->}[rrrr]^(.7){p_V'} & & &&B_4' \ar@{>>}[dd]_(.3){q_4'} \\
	&T\ar@{>->}[rrrr]|(.68)\hole_{x_1} \ar[dl]_{w} &&&& U\ar[dl]^(.4){u_1} && K_{m_3\circ q'_3} \ar@{>->}[rr]_-{k_{p'_V}} && K_{q'_4} \\
	Y \ar@{>->}[rrrr]_{y_1} & & && Q'_4}\]

By construction we have that 
\[m_3\circ q'_3 \circ \pi_{m_3 \circ q'_3}^1 = m_3\circ q'_3 \circ \pi_{m_3\circ q_3'}^2 \qquad  m_2\circ q_2' \circ \pi_{m_2 \circ q_2'}^1 = m_2\circ q'_2 \circ \pi_{m_2 \circ q_2'}^2 \]
but since $m_3$ and $m_2$ are monos this implies
\[q'_3 \circ \pi_{m_3 \circ q'_3}^1 = q'_3 \circ \pi_{m_3\circ q_3'}^2 \qquad q_2' \circ \pi_{m_2 \circ q_2'}^1 = q'_2 \circ \pi_{m_2 \circ q_2'}^2\]

Computing, we obtain
\[\begin{split}
	q \circ v_V \circ \pi_{q_4'}^1 \circ k_{p_V'} &= q \circ v_V \circ p_V' \circ \pi_{m_3 \circ q_3'}^1 \\
	&= q \circ w_V \circ \pi_{m_3 \circ q_3'}^1 \\
	&= w_Q \circ q_3' \circ \pi_{m_3 \circ q_3'}^1 \\
	&= w_Q \circ q_3' \circ \pi_{m_3 \circ q_3}^2 \\
	&= q \circ w_V \circ \pi_{m_3 \circ q_3'}^2 \\
	&= q \circ v_V \circ p_V' \circ \pi_{m_3 \circ q_3'}^2 \\
	&= q \circ v_V \circ \pi_{q'_4}^2 \circ k_{p'_V}
\end{split}\qquad\begin{split}
	q \circ v_V \circ \pi_{q'_4}^1 \circ k_{z_V'} &= q \circ v_V \circ z_V' \circ \pi_{m_2 \circ q_2'}^1 \\
	&= q \circ o_V \circ \pi_{m_2 \circ q_2'}^1 \\
	&= o_Q\circ q_2' \circ \pi_{m_2 \circ q_2'}^1 \\
	&= o_Q \circ q_2' \circ \pi_{m_2 \circ q_2}^2 \\
	&= q \circ o_V \circ \pi_{m_2 \circ q_2'}^2 \\
	&= q \circ v_V \circ t_V' \circ \pi_{m_2 \circ q_2'}^2 \\
	&= q \circ v_V \circ \pi_{q'_4}^2 \circ k_{z'_V}
\end{split}\]
Since the previous cube has a pushout as top face, by universal property, we have
\[
q \circ v_V \circ \pi_{q'_4}^1 = q \circ v_V \circ \pi_{q'_4}^2
\]

@@

\begin{lemma}\label{lemma:van_kampen}
	In $\EqHyp$, pushouts along regular monos are $\reg(\EqHyp)$-Van Kampen.
\end{lemma}

\begin{proof}
	In lieu of \Cref{lemma:stab}, it is enough to proof that, given a cube as the one below, with pullbacks as back faces, pushouts as bottom and top faces and such that $h$ is a regular mono,
	the front faces are pullbacks too, where $\mathcal{G}_i = (A_i, B_i, C_i, s_i, t_i, q_i)$, $\mathcal{G}'=(A_i', B_i', C_i', s_i', t_i', q_i')$, for $i = 1, 2, 3, 4$.
	\[
	\xymatrix@C=10pt@R=6pt{&\mathcal{G}_1'\ar[dd]|\hole_(.65){a}\ar[rr]^{h'} \ar[dl]_{k'} && \mathcal{G}_2' \ar[dd]^{b} \ar[dl]_{t'} \\ \mathcal{G}_3'  \ar[dd]_{c}\ar[rr]^(.7){p'} & & \mathcal{G}_4' \ar[dd]_(.3){d}\\&\mathcal{G}_1\ar[rr]|\hole^(.65){h} \ar[dl]^{k} && \mathcal{G}_2 \ar[dl]^{t} \\\mathcal{G}_3 \ar[rr]_{p} & & \mathcal{G}_4 }
	\]
	By \Cref{prop:eqhyp_complete} and \Cref{cor:mono_in_EqGrph}, the following two cubes have $\mathcal{M}$-pushouts as bottom faces and pullbacks as back faces,
	thus their front faces are pullbacks too.
	\[
	\xymatrix@C=10pt@R=6pt{&A_1'\ar[dd]|\hole_(.65){a_E}\ar[rr]^{h_E'} \ar[dl]_{k_E'} && A_2' \ar[dd]^{b_E} \ar[dl]_{t_E'} \\ A_3'  \ar[dd]_{c_E}\ar[rr]^(.7){p_E'} & & A_4' \ar[dd]_(.3){d_E}\\&A_1\ar[rr]|\hole^(.65){h_E} \ar[dl]^{k_E} && A_2 \ar[dl]^{t_E} \\A_3 \ar[rr]_{p_E} & & A_4 }
	\qquad
	\xymatrix@C=10pt@R=6pt{&B_1'\ar[dd]|\hole_(.65){a_V}\ar[rr]^{h_V'} \ar[dl]_{k_V'} && B_2' \ar[dd]^{b_V} \ar[dl]_{t_V'} \\ B_3'  \ar[dd]_{c_V}\ar[rr]^(.7){p_V'} & & B_4' \ar[dd]_(.3){d_V}\\&B_1\ar[rr]|\hole^(.65){h_V} \ar[dl]^{k_V} && B_2 \ar[dl]^{t_V} \\B_3 \ar[rr]_{p_V} & & B_4 }
	\]
	On the other hand we can consider the diagrams below, in which the inner squares are pullbacks.
	Since the outer diagrams commute, by definition of morphism of $\EqHyp$, then we have the existence of $m_2\colon C'_2\to U$, $m_3\colon C'_3\to Y $, $a_3\colon B'_3\to Y$ and $a_2\colon B'_2\to Y$.
	\[\xymatrix{C'_3 \ar@{.>}[dr]^{m_3} \ar@/^.3cm/[drr]^{p'_3} \ar@/_.3cm/[ddr]_{d_3} &&& C'_2 \ar@{.>}[dr]^{m_2} \ar@/^.3cm/[drr]^{t'_3} \ar@/_.3cm/[ddr]_{d_2}\\&Y \ar[r]_{y_1} \ar[d]_{y_2}& C'_4\ar[d]^{d_3}&& U \ar[d]_{u_2}\ar[r]_{u_1}& C'_4 \ar[d]^{d_3}\\&C_3 \ar[r]_{p_3}& C_4 &&C_2 \ar[r]_{t_3}& C_4 }\]

	\[\xymatrix{B'_3 \ar@{.>}[dr]^{a_3} \ar[r]^{p'_2} \ar[d]_{q'_3} &B'_4\ar[dr]^{q'_4}&& B'_2\ar@{.>}[dr]^{a_2} \ar[r]^{t'_2} \ar[d]_{q'_2} &B'_4\ar[dr]^{q'_4}\\C'_3 \ar[dr]_{d_3} &Y \ar[r]_{y_1} \ar[d]_{y_2}& C'_4\ar[d]^{d_3}&C_2' \ar[dr]_{d_2}& U \ar[d]_{u_2}\ar[r]_{u_1}& C'_4 \ar[d]^{d_3}\\&C_3 \ar[r]_{p_3}& C_4 &&C_2 \ar[r]_{t_3}& C_4 }\]

	Now, notice that  $m_3$ and $m_2$ are monos because $d_3$ and $d_2$ are regular monos. By the proof of \Cref{prop:eqhyp_complete}, to conclude it is enough to show that
	\[m_3\circ q'_3 = a_3 \qquad m_2\circ q'_2=a_2\]

	Indeed, if the previous equations hold, then $C'_3$ and $C'_2$ are epi-mono factorizations of $a_3$ and $a_2$ and the thesis follows from \Cref{cor:unique} and the proof of \Cref{prop:eqhyp_complete}.

	No if we compute we have:
	\[\begin{split}
		y_1\circ a_3&= q'_4\circ p'_2\\&=p'_3 \circ q'_3\\&=y_1\circ m_3\circ q'_3  
	\end{split}\qquad \begin{split}
		u_1\circ a_2&= q'_4\circ t'_2\\&=t'_3 \circ q'_3\\&=u_1\circ m_2\circ q'_2  
	\end{split}\]
	\[\begin{split}
		y_2\circ a_3&= d_3\circ q_3'\\&=y_2\circ m_3\circ q'_3
	\end{split}\qquad \begin{split}
		u_2\circ a_2&= d_2\circ q_2'\\&=u_2\circ m_2\circ q'_2
	\end{split}\]
	And we have done.
\end{proof}

@@

\section{Pros and cons of adhesive rewriting}
\label{rewriting}
The previous sections has shown how term graphs with equivalence,
possibly closed with respect to operator composition,
can be described as a suitable $\mathcal{M}$-adhesive category. 
The latter precisely capture EGGS, as originally
presented in~\cite{WillseyNWFTP21}. 
%
Being a purely graphical representation, EGGS allows for redundancy and possibly ambiguity,
so that e.g. a term such as $a \div a$ admits a few different representations as a term graph, as shown below
%
\[a \div a \mbox{ con 2 a sharate, con 2 a nella stessa classe di equivalenze, e con 2 a separate}\]

\begin{center}\begin{tikzpicture}
        \begin{pgfonlayer}{nodelayer}
                \node[style=small box](a) at (-1.8, 0){$a$};
                \node[style=none] (aatt) at (-1.5, 0){};
                \node[style=node](v) at (-0.9, 0){};
                \node[style=medium box](div) at (0, 0){$/$};
                \node[style=none](divfst) at (-0.3, -0.3){};
                \node[style=none] (divsnd) at (-0.3, 0.3){};
                \node[style=none] (divout) at (0.3, 0){};
                \node[style=node] (w) at (0.9, 0){};
        \end{pgfonlayer}
        \begin{pgfonlayer}{edgelayer}
                \draw(aatt.center) to (v);
                \draw[out=-60, in=-180] (v) to (divfst.center);
                \draw[out=60, in=-180](v) to (divsnd.center);
                \draw(divout.center) to (w);
        \end{pgfonlayer}
\end{tikzpicture}
\qquad
\begin{tikzpicture}
        \begin{pgfonlayer}{nodelayer}
                \node[style=small box](a1) at (-1.8, -0.5){$a$};
                \node[style=none] (a1att) at (-1.5, -0.5){};
                \node[style=node](v1) at (-0.9, -0.5){};
                \node[style=small box](a2) at (-1.8, 0.5){$a$};
                \node[style=none] (a2att) at (-1.5, 0.5){};
                \node[style=node](v2) at (-0.9, 0.5){};
                \node[style=medium box](div) at (0, 0){$/$};
                \node[style=none](divfst) at (-0.3, -0.3){};
                \node[style=none] (divsnd) at (-0.3, 0.3){};
                \node[style=none] (divout) at (0.3, 0){};
                \node[style=node] (w) at (0.9, 0){};
        \end{pgfonlayer}
        \begin{pgfonlayer}{edgelayer}
                \draw(a1att.center) to (v1);
                \draw(a2att.center) to (v2);
                \draw[out=0, in=-120] (v1) to (divfst.center);
                \draw[out=0, in=120](v2) to (divsnd.center);
                \draw(divout.center) to (w);
        \end{pgfonlayer}
        \begin{pgfonlayer}{eqlayer}
                \draw[dashed, rounded corners] (-1.2, -0.8) rectangle (-0.6, 0.8);
        \end{pgfonlayer}
\end{tikzpicture}
\qquad
\begin{tikzpicture}
        \begin{pgfonlayer}{nodelayer}
                \node[style=small box](a1) at (-1.8, -0.5){$a$};
                \node[style=none] (a1att) at (-1.5, -0.5){};
                \node[style=node](v1) at (-0.9, -0.5){};
                \node[style=small box](a2) at (-1.8, 0.5){$a$};
                \node[style=none] (a2att) at (-1.5, 0.5){};
                \node[style=node](v2) at (-0.9, 0.5){};
                \node[style=medium box](div) at (0, 0){$/$};
                \node[style=none](divfst) at (-0.3, -0.3){};
                \node[style=none] (divsnd) at (-0.3, 0.3){};
                \node[style=none] (divout) at (0.3, 0){};
                \node[style=node] (w) at (0.9, 0){};
        \end{pgfonlayer}
        \begin{pgfonlayer}{edgelayer}
                \draw(a1att.center) to (v1);
                \draw(a2att.center) to (v2);
                \draw[out=0, in=-120] (v1) to (divfst.center);
                \draw[out=0, in=120](v2) to (divsnd.center);
                \draw(divout.center) to (w);
        \end{pgfonlayer}
\end{tikzpicture}
\end{center}

%
Admittedly, the third one is illegal, i.e. it is an object of $\EqTGs$ but not one of $\GEqTGs$: constants have no input, 
and nodes that are targets of edges with the same label and source the empty word must be equivalent,
and possibly the same.
%
In general, note that both a maximally minimally and a shared representation of an EGG exist, and for our 
toy example are the left-most and the middle one above, respectively. In general terms,
for an EGG $\mathcal{G}$, such canonical forms are the final and the initial objects of the comma category 
$\mathcal{G}  \downarrow \GEqTGs$. 
The former part is a standard result in term graph rewriting, see e.g.~\cite{xxx}, while the latter is specific 
of the setting with equivalences.
%
The same construction carries on when restricting to the full category of $\GEqTGs$ of those objects whose 
underlying term graph is \emph{acyclic}.
%and in this case if $\mathcal{G}$ has no empty nodes, then the unique arrow 
%from $\mathcal{G}$ to the terminal object is easily shown to be injective on equivalence classes.

@@

As in the point above, uniqueness is guaranteed by faithfulness of $T$ and \Cref{rem:eqhyp_morphs}.	\qedhere 
	\end{enumerate}
\end{proof}

@@

\begin{proof}
\emph{Only if}

	\parbox{11cm}{	If $(h_E, h_V, h_Q)$ is mono, from \Cref{cor:mono1} we have that $h_E$ and $h_V$ are monos. 
	Let now $(f_E, f_V, f_Q), (g_E, g_V, g_Q): \mathcal{H\rightrightarrows K}$ be arrows equalized by $(h_E, h_V, h_Q)$. Let $e\colon E_Q\mto Q_{\mathcal{H}}$ be the equalizer in
	$\Set$ of $f_Q, g_Q\colon Q_{\mathcal{H}}\rightrightarrows Q_{\mathcal{K}}$. From the third point of \Cref{prop:fact} and \Cref{lim} we know that $h_V\colon V_{\mathcal{G}}\mto V_{\mathcal{H}} $ 
	is the equalizer of $f_V, g_V\colon V_{\mathcal{H}}\rightrightarrows V_{\mathcal{K}}$, thus by the second point of \Cref{prop:eqhyp_complete} we know that there is a mono 
	$m\colon Q_{\mathcal{G}}\mto E_Q$ fitting in the diagram aside. Now, notice that 
	$e\circ m\circ q_{\mathcal{G}}=q_\mathcal{H}\circ h_V=h_{Q}\circ q_{\mathcal{G}}$.
	Since $q_{\mathcal{G}}$ is epi we conclude that $e\circ m=h_Q$, from which the thesis follows.}
	\hfill
	\parbox{2cm}{\xymatrix@R=18pt{V_\mathcal{G} \ar@{>>}[d]_{q_{\mathcal{G}}} \ar@{>->}[r]^{h_V}& V_\mathcal{H} \ar@{>>}[dd]^{q_{\mathcal{H}}}\\Q_\mathcal{G} \ar@{>.>}[d]_{m}&\\ E_Q \ar@{>->}[r]_{e} & Q_\mathcal{{H}}}}

	\noindent
	\emph{If}
	Let us assume instead that $h_E$, $h_V$, and $h_Q$  are monos. XXXXX
\end{proof}

@@

>>>>

We can now prove that $\pbc$ is a class suitable to prove adhesivity with respect to it (a proof can be found in \Cref{proof:pbmono}).

\begin{restatable}{lemma}{pbm}\label{lem:pbmono}
	The class $\pbc$ contains all isomorphisms, it is closed under composition, decomposition and it is stable under pullbacks and pushouts.
\end{restatable}

We are now going to show that $\EqHyp$ is $\pbc$-adhesive.

@@

	\noindent 
	\parbox{8.5cm}{\hspace{15pt}
	For $Q'_1$, we can make a similar argument. Taking $(S, s_1, s_2)$ as the pullback of $m_2$ along $x_1$, we can use again the composition properties of pullbacks and the second point of \Cref{prop:eqhyp_complete} to guarantee the existence of a monomorphism $m_1\colon Q'_1\to S$ which makes the diagram aside commutative.}\hfill
	\parbox{4cm}{\xymatrix@R=15pt{B'_1\ar@{>>}[r]_{q'_1} \ar@{>->}[d]_{h'_V} \ar@/^.3cm/[rr]^{a_V} & Q'_1 \ar@{>.>}[d]_{m_1} & B_1 \ar@{>>}[dr]^{q_1}\\B'_2 \ar@{>>}[dr]_{q'_2} & S \ar@{>->}[r]^{s_2} \ar[d]_{s_1} & T \ar[r]^{x_2} \ar@{>->}[d]_{x_1}  & Q_1 \ar@{>->}[d]^{h_Q}\\& Q'_2 \ar@{>->}[r]_{m_2}& U \ar[r]_{u_2}& Q_2 }}

@@

\smallskip \noindent 
\parbox{8cm}{ \hspace{15pt}We want to extend such a morphism to one of $\EqHyp$ between $\mathcal{G}'_4\to \mathcal{H}$. If we are able to do so we can conclude because uniqueness is guaranteed by \Cref{rem:eqhyp_morphs}.

	\hspace{15pt}Now, consider the cube aside, $h$ is in $\pbc$ so by \Cref{lem:pb1} we know that its back face is a pullback. We can then apply the second point of \Cref{prop:kerset} to deduce that the square on the right is a pushout.

	By construction we have that 
	\begin{align*}
		m_3\circ q'_3 \circ \pi_{m_3 \circ q'_3}^1& = m_3\circ q'_3 \circ \pi_{m_3\circ q_3'}^2 \\
		m_2\circ q_2' \circ \pi_{m_2 \circ q_2'}^1 &= m_2\circ q'_2 \circ \pi_{m_2 \circ q_2'}^2
		\end{align*} 
	}\hfill\parbox{4cm}{\xymatrix@C=10pt@R=10pt{
	&B_1'\ar@{>>}[dd]|(.53)\hole_(.67){s_2\circ m_1\circ q_1'}\ar@{>->}[rr]^{h_V'} \ar[dl]_(.6){k_V'} && B_2' \ar@{>}[dd]^{m_2\circ q_2'} \ar[dl]_(.6){z_V'} & & \\
	B_3'  \ar[dd]_{m_3\circ q_3'}\ar@{>->}[rr]^(.7){p_V'}  &&B_4' \ar@{>>}[dd]_(.3){q_4'} \\
	&T\ar@{>->}[rr]_(.3){x_1}|(.5)\hole \ar[dl]_{w} && U\ar[dl]^(.4){u_1} \\
	Y \ar@{>->}[rr]_{y_1} & &  Q'_4\\K_{s_2\circ m_1\circ q'_1}  \ar@{>->}[rr]_-{k_{h'_V}} \ar[dd]_{k_{k'_V}}&& K_{m_2\circ q'_2} \ar[dd]^{k_{z'_V}}\\ \\ K_{m_3\circ q'_3} \ar@{>->}[rr]^-{k_{p'_V}} && K_{q'_4}}
}
	but since $m_3$ and $m_2$ are monos this implies
\[q'_3 \circ \pi_{m_3 \circ q'_3}^1 = q'_3 \circ \pi_{m_3\circ q_3'}^2 \qquad q_2' \circ \pi_{m_2 \circ q_2'}^1 = q'_2 \circ \pi_{m_2 \circ q_2'}^2\]

Computing, we obtain
\begin{align*}
	&q \circ v_V \circ \pi_{q_4'}^1 \circ k_{p_V'} = q \circ v_V \circ p_V' \circ \pi_{m_3 \circ q_3'}^1= q \circ w_V \circ \pi_{m_3 \circ q_3'}^1 = w_Q \circ q_3' \circ \pi_{m_3 \circ q_3'}^1 \\
	=& w_Q \circ q_3' \circ \pi_{m_3 \circ q_3}^2 = q \circ w_V \circ \pi_{m_3 \circ q_3'}^2 = q \circ v_V \circ p_V' \circ \pi_{m_3 \circ q_3'}^2 = q \circ v_V \circ \pi_{q'_4}^2 \circ k_{p'_V}\\
&q \circ v_V \circ \pi_{q'_4}^1 \circ k_{z_V'} = q \circ v_V \circ z_V' \circ \pi_{m_2 \circ q_2'}^1 = q \circ o_V \circ \pi_{m_2 \circ q_2'}^1 = o_Q\circ q_2' \circ \pi_{m_2 \circ q_2'}^1 \\
	= &o_Q \circ q_2' \circ \pi_{m_2 \circ q_2}^2 = q \circ o_V \circ \pi_{m_2 \circ q_2'}^2 = q \circ v_V \circ t_V' \circ \pi_{m_2 \circ q_2'}^2 = q \circ v_V \circ \pi_{q'_4}^2 \circ k_{z'_V}
\end{align*}

As already nopticed the square above is a pushout, hence we can conclude that
\[
q \circ v_V \circ \pi_{q'_4}^1 = q \circ v_V \circ \pi_{q'_4}^2
\]

@@

Having prove stability of $\pbc$-pushouts, we now turn to prove that they are Van Kampen with respect to regular monos.

\begin{lemma}\label{lemma:van_kampen}
	In $\EqHyp$, pushouts along arrows in $\pbc$ are $\reg(\EqHyp)$-Van Kampen.
\end{lemma}

\begin{proof}

Consider a cube as the one below on the left, with regular monos as vertical arrows,, with pullbacks as back faces, pushouts as bottom and top faces and such that $h$ is in $\pbc$.
In lieu of \Cref{lemma:stab}, if we show that 	the front faces are pullbacks too we can conclude.

To fix the notation, let $\mathcal{G}_i = (A_i, B_i, Q_i, s_i, t_i, q_i)$, $\mathcal{G}'=(A_i', B_i', Q_i', s_i', t_i', q_i')$, for $i = 1, 2, 3, 4$. By \Cref{prop:eqhyp_complete} and \Cref{cor:mono2} we know that the central and right cube below have pushouts as bottom faces and pullbacks as back faces,
	thus their front faces are pullbacks too.
	\[
	\xymatrix@C=10pt@R=10pt{&\mathcal{G}_1'\ar[dd]|\hole_(.65){a}\ar[rr]^{h'} \ar[dl]_{k'} && \mathcal{G}_2' \ar[dd]^{b} \ar[dl]_{z'} & &A_1'\ar@{>->}[dd]|\hole_(.65){a_E}\ar@{>->}[rr]^{h_E'} \ar[dl]_{k_E'} && A_2' \ar@{>->}[dd]_{b_E} \ar[dl]_{z_E'} &&B_1'\ar@{>->}[dd]|\hole_(.65){a_V}\ar@{>->}[rr]^{h_V'} \ar[dl]_{k_V'} && B_2' \ar@{>->}[dd]^{b_V} \ar[dl]_{z_V'} \\ 
		\mathcal{G}_3'  \ar[dd]_{c}\ar[rr]^(.7){p'} & & \mathcal{G}_4' \ar[dd]_(.3){d} &&A_3'  \ar@{>->}[dd]_{c_E}\ar@{>->}[rr]^(.7){p_E'} & & A_4' \ar@{>->}[dd]_(.3){d_E}
		&&B_3'  \ar@{>->}[dd]_{c_V}\ar@{>->}[rr]^(.7){p_V'} & & B_4' \ar@{>->}[dd]_(.3){d_V}\\
		&\mathcal{G}_1\ar[rr]|\hole^(.65){h} \ar[dl]^{k} && \mathcal{G}_2 \ar[dl]^{z} & &A_1\ar@{>->}[rr]|\hole^(.65){h_E} \ar[dl]^{k_E} && A_2 \ar[dl]^{z_E} &&B_1\ar@{>->}[rr]|\hole^(.65){h_V} \ar[dl]^{k_V} && B_2 \ar[dl]^{z_V} \\
		\mathcal{G}_3 \ar[rr]_{p} & & \mathcal{G}_4 && A_3 \ar@{>->}[rr]_{p_E} & & A_4&&B_3 \ar@{>->}[rr]_{p_V} & & B_4 }
	\]

Let us now consider the diagrams below, in which the inner squares are pullbacks.	Since the outer diagrams commute, by definition of morphism of $\EqHyp$, then we have the existence of $m_2\colon Q'_2\to U$, $m_3\colon Q'_3\to Y $, $a_3\colon B'_3\to Y$ and $a_2\colon B'_2\to Y$.

\smallskip 
\noindent 
\parbox{5.2cm}{\hspace{15pt}Now, notice that  $m_3$ and $m_2$ are monos because $c_Q$ and $b_2$ are injections. By the proof of \Cref{prop:eqhyp_complete}, to conclude it is enough to show that
	\[m_3\circ q'_3 = a_3 \qquad m_2\circ q'_2=a_2\]

		\hspace{15pt}Indeed, if the previous equations hold, then $Q'_3$ and   $Q'_2$ are images for $a_3$ and $a_2$ and the thesis follows from \Cref{prop:eqhyp_complete}.
	}\hfill \parbox{3.5cm}{\xymatrix@R=16pt{Q'_3 \ar@{.>}[r]_{m_3} \ar@{>->}@/^.3cm/[rr]^{p'_Q} \ar@{>->}@/_.3cm/[dr]_{c_Q} &Y \ar@{>->}[d]_{y_2}\ar@{>->}[r]_{y_1}& Q'_4 \ar@{>->}[d]^{d_Q}& Q'_2 \ar@{.>}[r]_{m_2} \ar@/^.3cm/[rr]^{z'_Q} \ar@{>->}@/_.3cm/[dr]_{b_Q}&U \ar[r]_{u_1} \ar@{>->}[d]_{u_2}& Q'_4\ar[d]^{d_Q}\\&Q_3 \ar@{>->}[r]_{p_3}& Q_4 &&Q_2 \ar[r]_{z_Q}& Q_4 }
\xymatrix@R=16pt{B'_3 \ar@{.>}[dr]^{a_3} \ar@{>->}[r]^{p'_V} \ar@{>>}[d]_{q'_3} &B'_4\ar@{>>}[dr]^{q'_4}&& B'_2\ar@{.>}[dr]^{a_2} \ar[r]^{z'_2} \ar@{>>}[d]_{q'_2} &B'_4\ar@{>>}[dr]^{q'_4}\\Q'_3 \ar@{>->}[dr]_{c_Q} &Y \ar@{>->}[r]_{y_1} \ar@{>->}[d]_{y_2}& Q'_4\ar[d]^{d_Q}&Q_2' \ar@{>->}[dr]_{c_Q}& U \ar@{>->}[d]_{u_2}\ar[r]_{u_1}& Q'_4 \ar@{>->}[d]^{d_Q}\\&Q_3 \ar@{>->}[r]_{p_Q}& Q_4 &&Q_2 \ar[r]_{z_Q}& Q_4 }}

	Now if we compute we have:
\[\begin{split}
	y_1\circ a_3&= q'_4\circ p'_2=p'_3 \circ q'_3=y_1\circ m_3\circ q'_3\\
u_1\circ a_2&= q'_4\circ t'_2=t'_3 \circ q'_3=u_1\circ m_2\circ q'_2 
\end{split} \qquad \begin{split}
 	y_2\circ a_3&= d_3\circ q_3'=y_2\circ m_3\circ q'_3\\
u_2\circ a_2&= d_2\circ q_2'=u_2\circ m_2\circ q'_2
\end{split}\]	
	And we have done.
\end{proof}

@@

\section{Pros and cons of adhesive rewriting}
\label{rewriting}
The previous sections has shown how term graphs with equivalence,
possibly closed with respect to operator composition,
can be described as a suitable $\mathcal{M}$-adhesive category. 
The latter precisely capture EGGS, as originally
presented in~\cite{WillseyNWFTP21}. 
%
Being a purely graphical representation, EGGS allows for redundancy and possibly ambiguity,
so that e.g. a term such as $a \div a$ admits a few different representations as a term graph, as shown below
%
\[a \div a \mbox{ con 2 a sharate, con 2 a nella stessa classe di equivalenze, e con 2 a separate}\]
%
Admittedly, the third one is illegal, i.e. it is an object of $\EqTGs$ but not one of $\GEqTGs$: constants have no input, 
and nodes that are targets of edges with the same label and source the empty word must be equivalent,
and possibly the same.
%
In general, note that both a maximally minimally and a shared representation of an EGG exist, and for our 
toy example are the left-most and the middle one above, respectively. In general terms,
for an EGG $\mathcal{G}$, such canonical forms are the final and the initial objects of the comma category 
$\mathcal{G}  \downarrow \GEqTGs$. 
The former part is a standard result in term graph rewriting, see e.g.~\cite{xxx}, while the latter is specific 
of the setting with equivalences.
%
The same construction carries on when restricting to the full category of $\GEqTGs$ of those objects whose 
underlying term graph is \emph{acyclic}.
%and in this case if $\mathcal{G}$ has no empty nodes, then the unique arrow 
%from $\mathcal{G}$ to the terminal object is easily shown to be injective on equivalence classes.

@@

As in the point aboce, uniqueness is guaranteed by faithfulness of $T$ and \Cref{rem:eqhyp_morphs}.	\qedhere 
	\end{enumerate}
\end{proof}

@@

\begin{proof}
	If $(h_E, h_V, h_Q)$ is mono, from \Cref{cor:mono1} we have that $h_E$ and $h_V$ are monos.

	\noindent 
	\parbox{2cm}{\xymatrix@R=18pt{V_\mathcal{G} \ar@{>>}[d]_{q_{\mathcal{G}}} \ar@{>->}[r]^{h_V}& V_\mathcal{H} \ar@{>>}[dd]^{q_{\mathcal{H}}}\\Q_\mathcal{G} \ar@{>.>}[d]_{m}&\\ E_Q \ar@{>->}[r]_{e} & Q_\mathcal{{H}}}}   
	\hfill 	\parbox{11cm}{Let now $(f_E, f_V, f_Q), (g_E, g_V, g_Q): \mathcal{H\rightrightarrows K}$ be arrows equalized by $(h_E, h_V, h_Q)$. Let $e\colon E_Q\mto Q_{\mathcal{H}}$ be the equalizer in $\Set$ of $f_Q, g_Q\colon Q_{\mathcal{H}}\rightrightarrows Q_{\mathcal{K}}$. From the third point of \Cref{prop:fact} and \Cref{lim} we know that $h_V\colon V_{\mathcal{G}}\mto V_{\mathcal{H}} $ is the equalizer of $f_V, g_V\colon V_{\mathcal{H}}\rightrightarrows V_{\mathcal{K}}$, thus by the second point of \Cref{prop:eqhyp_complete} we know that there is a mono $m\colon Q_{\mathcal{G}}\mto E_Q$ fitting in the diagram aside. Now, notice that 
	}\hfill 
	\[e\circ m\circ q_{\mathcal{G}}=q_\mathcal{H}\circ h_V=h_{Q}\circ q_{\mathcal{G}}\]
	Since $q_{\mathcal{G}}$ is epi we conclude that $e\circ m=h_Q$, from which the thesis follows.
\end{proof}

@@

\pbm*

\begin{proof}\label{proof:pbmono}
	Closure under composition and decomposition follows from \Cref{lem:pb1}, while stability under pushouts is the first point of \Cref{prop:kerset}. We are left with stability under pullbacks. 
	Let $h\colon \mathcal{G}_1\to \mathcal{G}_3$ be an arrow in $\pbc$ and $k\colon \mathcal{G}_2\to \mathcal{G}_3$ be another morphism of $\EqHyp$, with $\mathcal{G}_{i}=(E_i, V_i, Q_i, s_i, t_i, q_i)$.
